---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# talkbank

<!-- badges: start -->
[![R-CMD-check](https://github.com/rbcavanaugh/talkbank/workflows/R-CMD-check/badge.svg)](https://github.com/rbcavanaugh/talkbank/actions)
<!-- badges: end -->

The goal of talkbank is to provide easy access to talkbankDB through an R api. 

## Installation

You can install the released version of talkbank from [GitHub](https://github.com/) with:

```{r install}
# install.packages("devtools")
# devtools::install_github("rbcavanaugh/talkbank")
library(talkbank)
```

## Functionality

talkbank allows you to access data from talkbank.org through several functions. For example, you can access the utterances for a participant in aphasia bank

```{r example}
example_utterances <- getUtterances(corpusName = 'childes', corpora = c('childes', 'Eng-NA', 'MacWhinney', '010411a'));
example_utterances[1:5,]
```

The available functions for accessing different datatypes are:

```{r functions, eval=F}
getTranscripts()
getParticipants()
getTokens()
getUtterances()
getNgrams()
getTokenTypes()
getCQL()
```

These arguments for each function are documented, which can be accessed by typing ?functionName (e.g. ?getTranscriots()). Two helper functions are available to describe the arguments available for each function. getLegalValues() describes options available at each level of the database. validPath() will return whether or not a given path is valid, and where it is incorrect if applicable. 

```{r tree, eval=F}
getLegalValues()
```

Currently, talkpath only works for non-clinical banks at talkbank.org. For access to the clinical banks (e.g. aphaisaBank, dementiaBank), use talkbankDB.org. 

