---
title: "Example Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(kableExtra)
```

## A glimpse at each TBDBr query function

```{r setup}
library(TBDBr)
```

### getTranscripts

Queries a table where each row represents a transcript.

* Link to view transcript and play any associated media.
* Corpus path to transcript.
* Media types (audio/video) linked to transcript.
* Unique ID for transcript (PID).
* Languages spoken. Date recorded.
* Design Type.
* Activity Type.
* Group Type.

```{r transcripts}
# Get penglish-north american transcripts in in childes 
 transcripts <- getTranscripts(corpusName = 'childes',
                               corpora = c('childes', 'Eng-NA'))
```

```{r transcript output, echo = F}
kable(transcripts[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getParticipants

Queries a table where each row represents a participant (speaker) listed in a transcript. 

* Link to view transcript and play any associated media.
* Corpus path to transcript.
* Speaker's ID.
* Speaker's name.
* Speaker's role.
* Speaker's language.
* Speaker's age in months.
* Speaker's age in Years/Months/Days.
* Speaker's gender.
* Number of words spoken by speaker.
* Number of utterances spoken by speaker.
* Average number of words per speaker's utterance.
* Median number of words per speaker's utterance.

```{r participants}
# Get english-north american participants in childes 
 participants <- getParticipants(corpusName = 'childes',
                                 corpora = c('childes',
                                             'Eng-NA'))
```

```{r participant output, echo = F}
kable(participants[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getTokens

Queries a table with all the words from the selected transcripts, one word (token) per row.

* Link to view transcript and play any associated media.
* Corpus path to transcript.
* Utterance sequence number (starts at 0).
* Word sequence number within utterance (starts at 0).
* Speaker's role.
* Speaker's ID.
* The word (token).
* The word's stem.
* Part of speech code. (See CHAT manual for descriptions of codes).

```{r tokens}
# Get tokens (words) from one transcript.
tokens <- getTokens(corpusName = 'childes',
                    corpora = c('childes',
                                'Eng-NA',
                                'MacWhinney',
                                '010411a'));
```

```{r token output, echo = F}
kable(tokens[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getTokenTypes

Queries a table with all the words from the selected transcripts condensed into "types" based on word form and part of speech.

* Speaker's role.
* The word. Number of occurances of word in selected transcripts.
* Part of speech (See CHAT manual for descriptions of codes).
* The word's stem.

```{r tokentypes}
# Get token types from MacWhinney set.
token.types <- getTokenTypes(corpusName = 'childes',
                             corpora = c('childes',
                                         'Eng-NA',
                                         'MacWhinney'));
```

```{r tokentype output, echo = F}
kable(token.types[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getUtterances

Queries a table with all the words from the selected transcripts, one word (token) per row.

* Link to view transcript and play any associated media.
* Corpus path to transcript.
* Utterance sequence number (starts at 0).
* Word sequence number within utterance (starts at 0).
* Speaker's ID.
* Speaker's role.
* Utterance postcodes.
* Utterance GEMS.
* Utterance.
* Start time of utterance in associated media.
* End time of utterance in associated media.

```{r utterances}
utterances <- getUtterances(corpusName = 'childes',
                            corpora = c('childes',
                                        'Eng-NA',
                                        'MacWhinney',
                                        '010411a'))
```

```{r utterance output, echo = F}
kable(utterances[10:14,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getNgrams

Queries to get n-grams of specified size (n) and type.

* Speaker's role.
* The n-gram (word, stem, or part-of-speech). See CHAT manual for part-of-speech code values.
* Frequency count of n-gram.

```{r ngrams}
# Get 3-grams of words from one transcript.
ngrams <- getNgrams(nGram=c("3", "word"),
                    corpusName = 'childes',
                    corpora = c('childes',
                                'Eng-NA',
                                'MacWhinney',
                                '010411a'));
```

```{r ngram output, echo = F}
kable(ngrams[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```

### getCQL

Queryting by "CQL" (Corpus Query Language) lets us search for patterns in the selected transcripts. We construct a CQL query by specifying a search pattern of words, lemmas, and parts of speech. see documentation (?getCQL) for details. 

```{r cql}
# Query for text pattern "my ball" as lemma in MacWhinney set.
cql.myball <- getCQL(cqlArr=list(list(type="lemma", item="my", freq="once"),
                                 list(type="lemma", item="ball", freq="once")), 
                     corpusName = 'childes',
                     corpora = c('childes', 'Eng-NA', 'MacWhinney'));
```

```{r cql output, echo = F}
kable(cql.myball[1:5,]) %>% 
   kable_styling("striped") %>%
   scroll_box(width = "100%")
```
